A Recipe for Scavenging \textcolor{blue}{in Vertebrates} - the natural history of a behaviour

### 2016-07-13 response_letter.tex

\bigskip

We would like to thank the editor, subject editor and reviewers for their comments. By bringing attention to elements of the manuscript that require more careful communication we believe that you have helped us to significantly improve our manuscript. In this submission we have clarified these areas and added a body of work to bolster the conclusions drawn from our work. Please see below where we consider each of your points in turn.
For improving clarity in this document and in the manuscript, we highlighted our response in blue.

\section{Reviewer: 1}
This paper attempts to synthesize literature on traits that enable scavenging behavior in extant and extinct vertebrates. The paper is mostly well written, and most (but not all) pertinent literature sources are included. I do worry about the lack of novelty of the manuscript, and what I perceived as haphazard organization. Here are some specific issues that should be addressed in my opinion:

\begin{enumerate}
\item{Abstract: Scavenging can be difficult to observe, but probably no more so than predation. Also, using experimental carcasses, scavenging is actually quite easy to observe.}
\item{Abstract: You should point out that your ``scale of scavenging'' is only qualitative; you made no attempt to quantify such traits.}
\item{Introduction: Much of this section seems to follow DeVault's (2003) review paper on scavenging—in fact, there is almost nothing new here that is not covered in that earlier review.}

\textcolor{blue}{We rewrote the introduction to reflect more accurately the aim of this review.} %TG: i.e. the scavenging gradient allows us to infer scavenging in the fossil record?

\item{Introduction: You mention that you will use ``other methods in order to discern the most suitable morphologies...''  but you don't say what they are. What other methods are you talking about?}
\item{I think that early on in the paper you need to define obligate and facultative scavengers and then throughout the paper, specify which you are referring to. For example, in the first sentence of the Detection section, you talk about ``known scavengers'', but it is unclear if you really mean obligate scavengers. Nearly all carnivorous vertebrates are scavengers at least to some extent, so you need to elaborate on what you mean when you talk about ``scavengers''.}

%TG: maybe say something along the lines:
\textcolor{blue}{We do not consider a binary response to scavenging behavior (i.e. facultative/obligate) and argue that all carnivorous vertebrates are placed along a gradient of scavenging. We state this more clearly in the introduction: \textit{``there is no discrete divide between predators and scavengers but rather a continuous gradient (Pereira et al. 2014)''}.}

\item{Last sentence of Introduction: Again, you propose a ``scale of scavenging'', but without quantification, it is unclear how useful this ``scale'' is.}

\item{Page 6, 12: In your discussion of aquatic scavengers, you might want to refer to Beasley et al. (2012; Oikos 121:1021-1026).}

%TG: yep we did that

\item{Page 6: Why is jaw morphology likely to define scavengers?  Please elaborate.}

%TG: something along the lines: "The jaw is the principal tool for food ingestion and it's morphology will determine and the range of possible food items the organism is able to ingest"

\item{Page 9: It is true that carrion can comprise a large percentage of a hyena's diet, but your text seems misleading. Most studies have shown that most of a hyena's diet is from predation. For example, see Cooper et al. (1999; Afr. J. Ecol. 37:149–160) and Gasaway et al. (1991; Afr. J. Ecol. 29: 64–75).}
\item{Your section on Handling Time is so short that I think it should be cut or combined with another section.}

\textcolor{blue}{We rewrote the introduction to reflect more accurately the aim of this review.}

@@

\section{Reviewer: 3}
This is a well-written, enjoyable manuscript that assesses the functional traits associated to past and, mostly, current major vertebrate scavengers. As a result, the authors provide a sort of guide to identify the extent of scavenging behaviour among carnivores in both current and past ecosystems. I have several suggestions that should be considered before publication (please see below).

\subsection{Major comments:}

\begin{enumerate}
\item{Invertebrates are ignored completely throughout the ms. Please specify in the Introduction that this review focuses on vertebrate scavengers. In this line, I recommend to include ``in vertebrates'' at the end of the title to be more specific. Also, in the ``Competition'' section you mention competition with vertebrates and micro-organisms, but not with invertebrates. You should consider them in this section (and in ``Facilitation''; see next point).}

\textcolor{blue}{We changed the title of the manuscript to ``A Recipe for Scavenging in Vertebrates - the natural history of a behaviour''.}
%TG: Also maybe mention invertebrates in the competition part of the manuscript. + For a response to the reviewer, here is probably a good point for stating clearly that we wanted this manuscript to be focusing more on vertebrates because of the palaeo aspect. Both micro-organisms and invertebrates (I guess mainly arthropods) are unlikely to display preserved morphological clues for suggesting a scavenging behaviour because of their poor preservation anyway.

@@

\item{I found difficult to follow the main point of the paper because the ms is not well organized. Following a hierarchical order, I'd define four/five main sections: ``Introduction'', ``The Challenges of Scavenging'' (this could be a subsection within the previous section), ``Encounter Rate'' (with four different subsections: ``Metabolism'', ``Locomotion'', ``Sensory Detection'' and ``Carcass Availability''), ``Handling Time'' (which would include ``Food Processing'') and ``Conclusion''. \\

Moreover, ``Competition'' and ``Facilitation'' (I strongly encourage you to consider not only competitive, but also facilitative processes, which are mostly neglected in your review) should be treated as transversal factors that can modulate each of the abovementioned parameters. Thus, I'd include them in a different section, or mention the competitive and facilitative processes related to each parameter within each subsection. Fig. 1 should be re-organized accordingly.}

\textcolor{blue}{We clarified the structure of the manuscript by introducing clearer hierarchical subsections.}
%TG: + more blabalal on his ``Facilitation'' part

@@

\textcolor{blue}{The reference of Pobiner (2015) states that there were more individuals and we have changed the sentence to reflect this.}

\item{FOOD PROCESSING:} Page 18, 1st sentence: change ``enabling'' to ``enable''.

\textcolor{blue}{This has been changed.}

\item{FOOD PROCESSING:} Page 18, 3rd sentence: ``thus''? I don't understand this sentence; could you please reword it?

### 2016-07-13 response_letter.tex

\documentclass[12pt,letterpaper]{article}
\usepackage{url}
\usepackage{color}
%\signature{Your name}
%\address{Street \\ City \\ Country}
\begin{document}
%\begin{letter}{Company name \\ Street\\ City\\ Country}
To the editor

\bigskip

We would like to thank the editor, subject editor and reviewers for their comments. By bringing attention to elements of the manuscript that require more careful communication we believe that you have helped us to significantly improve our manuscript. In this submission we have clarified these areas and added a body of work to bolster the conclusions drawn from our work. Please see below where we consider each of your points in turn.
For improved clarity in this document and in the manuscript, we have highlighted our responses in \textcolor{blue}{blue}.

\section{Recommendation by the Subject Editor (anonymous):}
Three experienced reviews have evaluated your manuscript. Two of them are extremely critical of several aspects related to to the general handcraft and novelly of the review. The third reviewer is more positive, while making substantial recommendations for improvement of the manuscript. Should you be able to address all of the reviewers comments in a much revised version of the manuscript I would happy to reconsider the manuscript for consideration in Ecography and the E4 Award. This is not a provisional acceptance, since the manuscript would need to be revised again.

\smallskip

\textcolor{blue}{As we note below, we feel our earlier manuscript did not capture the hierarchy we had in mind. This updated version is more explicit in terms of the structure where we now highlight the sections and subsections. Regarding the novelty of our manuscript, we have changed our introduction which was regarded as too similar to deValult et al. (2003) and argue our case for what we consider inevitable overlap with Beasley et al. (2015).}

@@

\section{Reviewer: 1}
This paper attempts to synthesize literature on traits that enable scavenging behavior in extant and extinct vertebrates. The paper is mostly well written, and most (but not all) pertinent literature sources are included. I do worry about the lack of novelty of the manuscript, and what I perceived as haphazard organization. Here are some specific issues that should be addressed in my opinion:

\begin{enumerate}
\item{Abstract: Scavenging can be difficult to observe, but probably no more so than predation. Also, using experimental carcasses, scavenging is actually quite easy to observe.}
\item{Abstract: You should point out that your ``scale of scavenging'' is only qualitative; you made no attempt to quantify such traits.}
\item{Introduction: Much of this section seems to follow DeVault's (2003) review paper on scavenging—in fact, there is almost nothing new here that is not covered in that earlier review.}

\textcolor{blue}{We rewrote the introduction to more accurately reflect the aim of this review.} %TG: i.e. the scavenging gradient allows us to infer scavenging in the fossil record?

\item{Introduction: You mention that you will use ``other methods in order to discern the most suitable morphologies...''  but you don't say what they are. What other methods are you talking about?}

\textcolor{blue}{We now specify the approaches we have in mind:  \textit{``e.g. energetics models, comparative anatomy, palaeontology etc.'' }}

\item{I think that early on in the paper you need to define obligate and facultative scavengers and then throughout the paper, specify which you are referring to. For example, in the first sentence of the Detection section, you talk about ``known scavengers'', but it is unclear if you really mean obligate scavengers. Nearly all carnivorous vertebrates are scavengers at least to some extent, so you need to elaborate on what you mean when you talk about ``scavengers''.}

@@

%TG: maybe say something along the lines:
\textcolor{blue}{We do not consider a binary response to scavenging behavior (i.e. facultative/obligate) and argue that all carnivorous vertebrates are placed along a gradient of scavenging. We state this more clearly in the introduction: \textit{``there is no discrete divide between predators and scavengers but rather a continuous gradient (Pereira et al. 2014)''}.}

\item{Last sentence of Introduction: Again, you propose a ``scale of scavenging'', but without quantification, it is unclear how useful this ``scale'' is.}

\item{Page 6, 12: In your discussion of aquatic scavengers, you might want to refer to Beasley et al. (2012; Oikos 121:1021-1026).}

\textcolor{blue}{We have included this reference and explicitly discuss how differences in the physical properties of water Vs air can affect scavenging: \textit{``The physical differences between water and air mean carcass availability is radically different between these environments (Beasley et al. 2012). For one, carcasses get moved around by water which results in a more diffuse signal being produced for would-be scavengers. Carcasses also tend to sink in water where they are no longer accessible to pelagic scavengers (Beasley et al. 2012). Research has shown that an animal need only travel between 16 to 36 km to encounter a fresh whale carcass  (Moleón and Sánchez-Zapata 2015). The phenomenon of occasional bounties of carrion in the form of these whale falls has led some researchers to investigate if an obligate scavenger could survive by seeking out these remains exclusively.''}}

@@

\item{Page 6: Why is jaw morphology likely to define scavengers?  Please elaborate.}

%TG: something along the lines: "The jaw is the principal tool for food ingestion and it's morphology will determine and the range of possible food items the organism is able to ingest"

\item{Page 9: It is true that carrion can comprise a large percentage of a hyena's diet, but your text seems misleading. Most studies have shown that most of a hyena's diet is from predation. For example, see Cooper et al. (1999; Afr. J. Ecol. 37:149–160) and Gasaway et al. (1991; Afr. J. Ecol. 29: 64–75).}

\textcolor{blue}{We have clarified this point about hyenas to say it is their ability to take large proportions of carrion under certain circumstances that makes them a useful focal species: \textit{``The particular ability of hyenas to subsist on high proportions of carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms.''}}

\item{Your section on Handling Time is so short that I think it should be cut or combined with another section.}

\textcolor{blue}{The hierarchy of our paper wasn't clear in the first version but we have now clarified that this is the lead in to a larger section which has subsections e.g. food processing}

@@

\end{enumerate}

\section{Reviewer: 2}
This article is a review of the literature what the authors term a ``natural history'' of scavenging. They have approached the topic from the perspective of ecological variables (and other variables that condition them) in optimal foraging theory including encounter rate and handling time, but also metabolism, locomotion (e.g soaring and swimming), detection, prey availability, competition, and food processing. They describe many useful examples of the behavior and ecology of obligate and facultative scavengers. But I think they often go to far in their assumptions in trying to weave these examples together in a synthetic way. Some examples are inferring scrounging behaviors in flying pterosaurs (pace 11), assuming that hominins used vultures to identify scavenging opportunities (page 11), surmising that scavenging evolved with the earliest tetrapods based on ``potential carrion unexploited by marine vertebrates'' (page 12). There is a lot of ``these researchers have suggested X'' and then taking that as a given fact.

\smallskip
\textcolor{blue}{One of the central points of our paper is to compare the biology of extant scavengers with extinct species in order to determine if they were suited to scavenging. All of the examples we use are backed up by references from previous work. Rather than taking them as fact we offer these as possible behaviours that the extinct species engaged in. Because we can't observe their behaviour directly we are left to make assumptions about how likely they were to have enganged in them. But this is hardly unique to our work. We have added some extra references in support of our claims: \textit {``Thus, given pterosaurs seem to have cohabited in large numbers (Witton 2013), and the known benefits this can have for social foraging in birds (Dermody et al. 2011), it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.''}}
\smallskip

While the authors distinguish between obligate and facultative scavenging, they tend to lump them together when it comes to constructing their arguments. In the end, is the argument about the former, or the latter? Is the real interest in whether species scavenge at all, what proportion of their resources are obtained by scavenging, or what the conditions are that produce facultative or obligate scavengers?

They also seem to ignore flexibility in species' behaviors. For instance, while spotted hyenas are on their highly scavenging side of their Figures 1 and 2, in some ecosystems they hunt significant proportion of their prey. Similarly, lions can take a large proportion of their prey as scavenged, as the authors note. But to suggest, as they do in their abstract, that they can apply their scale of scavenging to any species at any time to judge how important scavenging was in its diet is frankly completely ignoring behavioral variability over time and space. 
\smallskip

\textcolor{blue}{It is certainly true that facultative scavengers can be variable in the proportion of carrion they consume but we would argue that species like hyenas have certain characteristics that make this possible. In contrast, cheetahs do not have a biology that lends itself to scavenging in the wild, due to poor competitive ability etc. We appreciate that there is a continuum across and among species which we specifically note in this revised version.} 
\smallskip

It is substantially more difficult with fossil species for which we may know far fewer of the variables they list in Figure 1, in a nuanced way. Figure 2 does not display the diversity of scavengers through time, it displays examples of species that display observed or inferred scavenging behaviors. Diversity has specific ecological definition(s) that are being ignored here.

@@

If the authors chose to resubmit, I would suggest they organize their manuscript based on known information about modern species, known (e.g. bone processing by theropods) and inferred (e.g. scavenging in hominins) information about fossil species, and then try to draw broader conclusions. They might look to articles about other modern scavenging opportunities such as Capaldo and Peter's study of wildebeest drownings \url{http://www.sciencedirect.com/science/article/pii/S0305440385700398}, and suggestions that sabertooth cats which they claim were unable to feed on bones may have consumed more bone than penecontemporaneous dire wolves \url{https://www.jstor.org/stable/20627167?seq=1#page_scan_tab_contents}.

In sum, while I think the aim of the article is a very worthwhile one, I think the authors cannot claim to have produced an all-encompassing model that can predict the presence or level of scavenging in any organism - because I think this simply cannot be done.

Finally, some of the perspectives of this manuscript overlap substantially with parts of Beasley et al. 2015 (pages 108-112), which is cited numerous times in the manuscript.
\smallskip

\textcolor{blue}{The sections of Beasley et al. (2015) are the evolution of scavengers, competition between scavengers and environmental effects on scavenger ecology. These are very broad areas so in a review of this sort some overlap is inevitable. We feel our emphasis on extinct species and systems as well as the frame of optimal foraging theory gives this manuscript its novelty.} 
\smallskip

@@

\item{PREY AVAILABILITY:} Page 12, 2nd paragraph, 1st sentence: as said before, this is not true, given that scavenging may indirectly affect herbivore populations and thus carrion availability (see Mole\'{o}n et al. 2014 Biol. Rev.).

\textcolor{blue}{We have removed the part of the sentence where we mistakenly suggest that scavengers cannot affect carrion availability.}

\item{PREY AVAILABILITY:} Page 13, 3rd and 4th sentences: whaling has exerted also a huge negative impact on scavengers by depleting whale populations. If you are going to talk about human impacts, maybe is better to do it at the end of the ms, perhaps in a new section in which you explore which scavenging traits could be more successful in a human dominated world.

\textcolor{blue}{We have decided to remove this section because we felt human impacts go beyond the scope of our paper.}

### 2016-07-13 natural history of scavenging.tex
added facilitation sections

Whatever the case, it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
Today, terrestrial scavenging in mammals is probably best known in the African Savannah context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet \textcolor{blue}{\citep{pereira2014facultative,periquet2015lion}.}
In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \textcolor{blue}{\citep{devault2003scavenging,beasley2015vertebrates,pereira2014facultative}}.
The particular reliance of hyenas on carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms \textcolor{blue}{\citep{periquet2015lion}}. 
In terms of locomotion, they employ a characteristic ``rocking horse gait''  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
Such long-distance travel is apparent \textcolor{blue}{in many other cursorial predators \citep{pennycuick1995radius,janis2014forelimb,pereira2014facultative}}. 
In contrast, big cats rely on ambush \citep{pennycuick1995radius,pereira2014facultative}. 
This difference means ambush predators tend to be more obligate hunters \citep{pereira2014facultative} 
These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 - 5.3 Mya), were likely to have been ambush predators which would argue against them taking a lot of carrion.

@@

Existing in this environment also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \textcolor{blue}{\citep{jackson2008effect,KaneVul,moleon2014inter}}.
These interactions are properly three dimensional in the sense of \cite{pawar2012dimensionality} because both producer and scrounger are in the air \citep{dall2005information}. 
Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
These behaviours allow them to forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \textcolor{blue}{\citep{cortes2014bird}}.
Information transfer of this kind is typically inadvertent and as a consequence no complex social interactions are required, simply the ability to recognise a successful forager.
Thus, it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.

@@

\subsection{\textcolor{blue}{Carcass} Availability}
The environmental influence on \textcolor{blue}{carcass} availability is an aspect that greatly affects encounter rate. %but is invisible to the selective forces acting on the scavenger. 
Aspects including, primary productivity, relief, and temperature will all affect scavenging tendency. 
\cite{ruxton2004obligate} suggest an historic ecosystem with a productivity similar to the Serengeti could have supported an \textit{obligate} mammalian or reptilian terrestrial scavenger.
Indeed, in systems that were dominated by large ectothermic or mesothermic herbivore vertebrates, the same primary productivity would have supported a greater biomass, due to the scaling of mass with metabolic rate \citep{mcnab2009resources}.
The upshot of this may have been a higher biomass of herbivores dying and offering scavenging opportunities (although these larger species may have also lived longer).

@@

Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
The geological record shows the Earth has undergone radical fluctuations in temperature over time.
This will have had a significant bearing on the availability and persistence of carrion.
To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was on average at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
Moreover, oceanic productivity and habitat structure are all impacted by climactic conditions.
The impacts these can have on scavengers have been empirically supported e.g. \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".

@@

In addition to fending off other vertebrates, scavengers also have to contend with competition from micro-organisms, which requires a specialised physiology. 
Although the findings of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appear sound, the fact remains that overcoming micro-organism toxins is still a beneficial adaptation to any scavenger. 
Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
This adaptation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans}; so-called ``vultures of the seas'') had an average pH of 1.5, which enables them to consume fisheries discards and squid carcasses. 
There is also evidence of selection for ``toxification minimizers'' beyond birds among the ectotherms.
From our earlier arguments we know that ecthotherms are limited in their ability to find carrion as quickly as endotherms. 
These later arrivers would thus benefit especially from well-developed detoxifying apparatus. 
\cite{shivik2006vultures} suggests that ``specialized oral structures in snakes may have evolved under pressures associated with scavenging."
Moreover, some researchers have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}. 
\textcolor{blue}{In water, by contrast, competition with micro-organims is significantly reduced because of the high pressures and low temperatures \citep{beasley2012carrion}.}

@@

Again, we can draw a comparison with clades that are lacking in these features. 
Extant bats are poorly equipped when it comes to feeding on carrion; the larger forms are typically frugivores and therefore lack the adaptations for digesting meat, while the smaller carnivorous bats are mainly found in the microbats, which are insectivorous \citep{aguirre2003implications}.   
That said, Necromantis (“death-eater”), a large bat from the middle to late Eocene (56 - 33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for an extinct scavenging bat \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.

@@

>>>>

Whatever the case, it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range more widely, a vital component in seeking out carrion.
Modern endothermic mammals can sustain longer periods of energetically expensive activity \citep{bennett1979endothermy} resulting in larger foraging ranges. 
Today, terrestrial scavenging in mammals is probably best known in the African Savannah context where hyenas, jackals and lions all take sizeable proportions of carrion in their diet \textcolor{blue}{\citep{pereira2014facultative,periquet2015lion}.}
In the spotted hyena (\textit{Crocuta crocuta}), striped hyena (\textit{Hyaena hyaena}) and brown hyena (\textit{Hyaena brunnea}) it can be over 90\% \citep{jones2015african}.
And although no contemporary terrestrial vertebrate exists as an obligate scavenger, most, if not all, are facultative to some extent \textcolor{blue}{\citep{devault2003scavenging,beasley2015vertebrates,pereira2014facultative}}.
\textcolor{blue}{The particular ability of hyenas to subsist on high proportions of carrion means we can use them as examples of efficient terrestrial scavengers to compare with other forms \citep{periquet2015lion}}. 
In terms of locomotion, they employ a characteristic ``rocking horse gait''  which allows them to cover great distances efficiently, loping at 10 km/hr \citep{mills1989comparative,jones2015african}. 
Such long-distance travel is apparent \textcolor{blue}{in many other cursorial predators \citep{pennycuick1995radius,janis2014forelimb,pereira2014facultative}. 
In contrast, big cats rely on ambush \citep{pennycuick1995radius,pereira2014facultative}. 
This difference means ambush predators tend to be more obligate hunters \citep{pereira2014facultative}.} 
These insights allow us to compare extant terrestrial species to their prehistoric forebears given the dominance of mammalian carnviores since the Eocene (56-33.9 Mya) where the order split into the Caniforma and Feliforma \citep{van1987skeletal}.
To take one example, \cite{anyonge1996locomotor} found that \textit{Nimravides}, a genus of sabretooth cat from the Miocence (10.3 - 5.3 Mya), were likely to have been ambush predators which would argue against them taking a lot of carrion.

@@

Existing in this environment also means being able to gather a wealth of information from other foragers, be they conspecifics or otherwise \textcolor{blue}{\citep{jackson2008effect,KaneVul,moleon2014inter}}.
These interactions are properly three dimensional in the sense of \cite{pawar2012dimensionality} because both producer and scrounger are in the air \citep{dall2005information}. 
Again, returning to vultures, the genus \textit{Gyps} consists of highly social and colonially nesting species \citep{fernandez2015density}.
These behaviours allow them to forage far more efficiently because one bird can scrounge information on the location of food from another successful forager \textcolor{blue}{\citep{cortes2014bird}}.
Information transfer of this kind is typically inadvertent and as a consequence no complex social interactions are required, simply the ability to recognise a successful forager.
\textcolor{blue}{Thus, given pterosaurs seem to have cohabited in large numbers \citep{witton2013pterosaurs}, and the known benefits this can have for social foraging in birds \citep{jackson2011evolutionary}, it seems probable that scrounging behaviours were seen in the flying pterosaurs as well.}

@@

\subsection{\textcolor{blue}{Abiotic effects on Carcass} Availability}
The environmental influence on \textcolor{blue}{carcass} availability is an aspect that greatly affects encounter rate. %but is invisible to the selective forces acting on the scavenger. 
Aspects including, primary productivity, relief, and temperature will all affect scavenging tendency.

\cite{ruxton2004obligate} suggest an historic ecosystem with a productivity similar to the Serengeti could have supported an \textit{obligate} mammalian or reptilian terrestrial scavenger.
Indeed, in systems that were dominated by large ectothermic or mesothermic herbivore vertebrates, the same primary productivity would have supported a greater biomass, due to the scaling of mass with metabolic rate \citep{mcnab2009resources}.
The upshot of this may have been a higher biomass of herbivores dying and offering scavenging opportunities (although these larger species may have also lived longer).

@@

Perhaps the greatest environmental driver of scavenging tendency is that of temperature. 
his has a significant bearing on the availability and persistence of carrion.
We know from the geological record shows that the Earth has undergone radical fluctuations in temperature over time.
\textcolor{blue}{TOn land, there are a wide variety of vegetation types, from thickly forested areas to open grassland and the extent of this vegetation changes with season.}
All of these factors influence the availabiltiy of carrion.  
Moreover, oceanic productivity is also impacted by climactic conditions.

To illustrate the point, a 10$^{\circ}$C increase in ambient temperature can double carcass decomposition rates \citep{parmenter2009carrion} and geological evidence indicates that the Mesozoic Earth was on average at least 6 $^{\circ}$C warmer than now \citep{sellwood2006mesozoic}.
In terms of specific habitats, it has been shown that decomposition is greater in warm and moist areas versus more xeric ones \citep{beasley2015vertebrates}.
The impacts these can have on scavengers have been empirically supported e.g. \cite{beasley2015vertebrates} who point to a series of studies showing how microbes and invertebrates benefit at higher temperatures to the detriment of vertebrate scavengers such that ``above 20$^{\circ}$C vertebrates were able to detect and consume only 19\% of small-mammal carcasses, whereas at temperatures below 18$^{\circ}$C, vertebrates consumed 49\% of carcasses".
\textcolor{blue}{In savannah, dense vegetation can prevent vultures from landing to a carcass because they will be unable to take off again \citep{bamford2009effect}. 
The Earth has also undergone a series of ice ages of various spatio-temporal extent \citep{diedrich2012cave} and modern terrestrial settings that experience sub-zero conditions can act as a microcosm to show the effect extreme cold can have on scavenging.
Notably, freezing carcasses can become too hard to consume by most vertebrate carnviores \citep{selva2003scavenging}.}

\subsection{\textcolor{blue}{Biotic effects on Carcass} Availability}
Through facilitation e.g. local enhancement, adept scavengers may increase the amount of carrion mammalian carnivores feed on thus increasing their population size \citep{moleon2014inter}. 
This higher population of predators will take more prey and so produce more carcasses. Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.

@@

In addition to fending off other vertebrates, scavengers also have to contend with competition from micro-organisms, which requires a specialised physiology. 
Although the findings of \cite{shivik2006vultures} that ``evolutionary pressures favor detection maximizers relative to toxification minimizers in competitive interactions for carcasses." appear sound, the fact remains that overcoming micro-organism toxins is still a beneficial adaptation to any scavenger. 
Avian scavengers have evolved incredibly acidic stomachs that allow them to consume and process putrefied flesh with no ill effects \citep{houston1975digestive,roggenbuck2014microbiome}. 
This adaptation is not restricted to vultures though, \cite{gremillet2012vultures} showed wandering albatrosses (\textit{Diomedea exulans}; so-called ``vultures of the seas'') had an average pH of 1.5, which enables them to consume fisheries discards and squid carcasses. 
There is also evidence of selection for ``toxification minimizers'' beyond birds among the ectotherms.
From our earlier arguments we know that ecthotherms are limited in their ability to find carrion as quickly as endotherms. 
These later arrivers would thus benefit especially from well-developed detoxifying apparatus. 
\cite{shivik2006vultures} suggests that ``specialized oral structures in snakes may have evolved under pressures associated with scavenging."
Moreover, some researchers have charted an evolutionary course from basal fossorial snakes to modern terrestrial species by way of an obligate scavenger intermediate \citep{bauchot2006snakes}. 
\textcolor{blue}{In water, by contrast, competition with micro-organisms is significantly reduced because of the high pressures and low temperatures \citep{beasley2012carrion}.}

\subsection{Facilitation}
As well as competitive interactions, there are also facilitatory processes at play between vertebrate scavengers that can reduce handling time for the facilitated species. 
Some examples include vultures, hyenas and wolves opening the tough hides of ungulate carcasses that would otherwise be inaccessible to corvids and smaller mammalian carnivores \citep{selva2003scavenging,moleon2015carcass}.

@@

Again, we can draw a comparison with species that are lacking in these features. 
\textcolor{blue}{Despite readily eating carrion in captivity, cheetahs (\textit{Acinonyx jubatus}) and African Wild Dogs (\textit{Lycaon pictus}) rarely do so in the wild because they are subordinate to many of the mammalian competitors they coexist with \citep{pereira2014facultative}.} 
Extant bats are poorly equipped when it comes to feeding on carrion; the larger forms are typically frugivores and therefore lack the adaptations for digesting meat, while the smaller carnivorous bats are mainly found in the microbats, which are insectivorous \citep{aguirre2003implications}.   
That said, Necromantis (“death-eater”), a large bat from the middle to late Eocene (56 - 33.9 Mya) had a robust cranio-mandibular morphology, and is a likely candidate for an extinct scavenging bat \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.

@@

% Modern scavenging assemblages can be inflenced by carcass size \citep{moleon2015carcass}. 
% Larger carcasses tend to last longer and also present a more conspicuous target for a foraging scavenger which results in more species attending them \citep{moleon2015carcass}.
% This will have had implications for extinct assemblages because body size distribtions vary across different environments but also across time. 
% \cite{10.1371/journal.pone.0051925} showed Mesozoic faunal distributions may have been skewed towards larger species even when fossil biases are taken into account. 
% If this pattern is real then the scavenger assemblages during this era would have been particularly diverse.

% As well as competitive interactions, there are also facilitatory processes at play between vertebrate scavengers that can reduce handling time. 
% Rather than there being a random assortment of species at a carcass, scavenging assemblages tend to be nested \citep{selva2007nested}. 
% This means the species that feed on the majority of carcasses, for example vultures, are a subset of a larger community which comprises hyenas, jackals, raptors etc. \citep{sebastian2016nested}. 
% In extant systems these interactions can reduce handling times for the facilitated species. 
% Some examples include vultures, hyenas and wolves opening the tough hides of ungulate carcasses that would otherwise be inaccessible to corvids and smaller mammalian carnivores \citep{selva2003scavenging,moleon2015carcass}. 
% The availability of carcasses can be indirectly influenced by the action of scavengers. 
% Through facilitation e.g. local enhancement, vultures may increase the amount of carrion mammalian carnivores feed on thus increasing their population size \citep{moleon2014inter}. 
% This higher population of predators will take more prey and so produce more carcasses. Conversely, by feeding on carcasses, predators may hunt less because they are sated by carrion, ultimately reducing predation risk on their prey base \citep{moleon2014inter}.
